---
title: "EXCEL"
author: "Jay Brophy"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  html_document:
    code_folding: hide
    highlight: tango
    theme: cerulean
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
link-citations: yes
linkcolor: red
urlcolor: blue
---


```{r setup, include=FALSE, results='asis', warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages <- c("ggplot2", "metafor", "meta", "tidyverse", "rjags", "rstan", "R2jags", "bayesmeta", "mcmcplots", "ggthemes", "knitr","kableExtra", "coda")
invisible(lapply(packages, function(xxx) suppressMessages(require(xxx, character.only = TRUE, quietly=TRUE, warn.conflicts = FALSE))))
```

## Produces EXCEL figures for CJC Bayesian publication

### Entering data primary outcome

```{r results='asis', warning=FALSE, message=FALSE}

# mortality data entered as NOBLE, SYNTAX,  EXCEL, PRECOMBAT (leave oout these small trials Boudriot) 
# composite endpoint for Noble = sum mortality+Mi+stroke PCI= 54+43+21=118; CABG=50+15+12=77
Mc<- 89 #mortality in non-expose (CABG)
Nc<-957 #total in non-expose
Me<-119 #mortality in expose (PCI)
Ne<- 948 #total in expose
Pc<-176 #primary composite Death, MI, stroke  in control in non-expose (CABG)
Pe<-203 #primary composite Death, MI, stroke  in control in in expose (PCI)
Re <- 290 #event in expose (PCI) - MACCE includes revasc
Rc <- 228 #event in non-expose (CABG) - MACCE includes revasc


temp <- data.frame(Mc=Mc,Me=Me,Pc=Pc, Pe=Pe, Rc=Rc, Re=Re, Nc=Nc, Ne=Ne)
temp$Study <- c("EXCEL")
temp <- temp[c(9,1:8)] # reorder columns

kable(temp, caption="Outcomes at 5 years") %>%
  kable_styling(bootstrap_options = "striped", full_width = F)

 #write.csv(temp, file = "data1.csv", row.names = FALSE)
Ex_p_c <- temp[1,4]
Ex_p_e <- temp[1,5]
Ex_n_c <- temp[1,8]
Ex_n_e <- temp[1,9]

set.seed(1234)
#Prior is beta(1,1)
# sampling 100,000 random variables from posterior
post_Ex_p_c <- rbeta(100000, Ex_p_c + 1, Ex_n_c - Ex_p_c + 1 )
post_Ex_p_e <- rbeta(100000, Ex_p_e + 1, Ex_n_e - Ex_p_e + 1 )

# calculting posterior of differences
post_Ex_p_diff <- post_Ex_p_e - post_Ex_p_c
paste("EXCEL data alone - Differences in MACE between PCI & CABG")
quantile(post_Ex_p_diff, probs = c(0.025, .5, 0.975))

# probabilities >0 and >1
paste("EXCEL data alone - Probability PCI worse than CABG = ", sum(post_Ex_p_diff*100 >0)/100000)
paste("EXCEL data alone - Probability PCI worse than CABG by >1% = ", sum(post_Ex_p_diff*100 >1)/100000)


# given large sample sizes, can verify answers with normal approximation
paste("With normal approximation, EXCEL data alone - Probability PCI worse than CABG = ", round(1-pnorm(0, mean(post_Ex_p_diff*100), sd(post_Ex_p_diff*100)),3))
paste("With normal approximation, EXCEL data alone - Probability PCI worse than CABG > 1% = ", round(1-pnorm(1, mean(post_Ex_p_diff*100), sd(post_Ex_p_diff*100)),3))

```

The calculated differrence in event ratios are based on the ratios of events to total populations in each arm. They do not include time to event as reported in the orginal NEJM publication and are consequentlty slightly different than the reported event rates.

### EXCEL Primary outcome graph - Figure 1a

```{r}
fig1a <- ggplot(data.frame(x = c(-3, 8)), aes(x = x)) +
        stat_function(fun = dnorm, args = list(mean(post_Ex_p_diff*100), sd(post_Ex_p_diff*100)),  colour = "orange", size=2) +
        scale_x_continuous(name = "Primary outcome difference (PCI - CABG) -> PCI worse") +
        scale_y_continuous(name = "Density") +
        labs(title="EXCEL PDF for primary composite outcome (risk difference)",
             subtitle = "Non-informative prior", caption="AUC = area under the curve\n     PDF = probability density function") +
        geom_vline(xintercept=mean(post_Ex_p_diff*100)) +
        annotate("text", label = "Black vertical line =\nmean outcome difference \n(3.0%) increased with PCI", x = 5.0, y = .2, color = "black", hjust = 0) +
        annotate("text", label = "Grey AUC = \nprobability (87%) \nPCI > CABG outcome \nby > 1%", x = 5.5, y = .12, color = "black", hjust = 0) +
         annotate("text", label = "Grey + yellow AUC =\nprobability (95%) \nPCI > CABG outcome", x = -2, y = .18, color = "black", hjust = 0) +
        theme_classic(base_size = 10) +
        stat_function(fun = dnorm, args = list(mean(post_Ex_p_diff*100), sd(post_Ex_p_diff*100)), xlim = c(1,8), geom = "area", alpha = 0.2) +
        stat_function(fun = dnorm, args = list(mean(post_Ex_p_diff*100), sd(post_Ex_p_diff*100)), xlim = c(0,1), geom = "area", alpha = 0.2, fill = "yellow") 
fig1a

pdf("output/fig1a.pdf")
fig1a 
dev.off()

ggsave("output/fig1a.png")  
```


### EXCEL Mortality difference - Figure 1b
Bayesian probability difference based on non-informative prior

```{r}
# Excel data
Ex_m_c <- temp[1,2]
Ex_m_e <- temp[1,3]
Ex_n_c <- temp[1,8]
Ex_n_e <- temp[1,9]

#Prior is beta(1,1)
# sampling 10,000 RV from posterior
set.seed(1234)
post_Ex_m_c <- rbeta(10000, Ex_m_c + 1, Ex_n_c - Ex_m_c + 1 )
post_Ex_m_e <- rbeta(10000, Ex_m_e + 1, Ex_n_e - Ex_m_e + 1 )

# calculting posterior of differences
post_Ex_m_diff <- post_Ex_m_e - post_Ex_m_c
quantile(post_Ex_m_diff, probs = c(0.025, .5, 0.975))

1-pnorm(0, mean(post_Ex_m_diff*100), sd(post_Ex_m_diff*100))
1-pnorm(1, mean(post_Ex_m_diff*100), sd(post_Ex_m_diff*100))

# graph
fig1b <- ggplot(data.frame(x = c(-2, 7.5)), aes(x = x)) +
        stat_function(fun = dnorm, args = list(mean(post_Ex_m_diff*100), sd(post_Ex_m_diff*100)), colour = "orange", size=2) +
        scale_x_continuous(name = "Total mortality difference (PCI - CABG) -> PCI worse") +
        scale_y_continuous(name = "Density") +
        labs(title="EXCEL PDF for total mortality (risk difference)",
             subtitle = "Non-informative prior", caption="AUC = area under the curve\n     PDF = probability density function") +
        geom_vline(xintercept=mean(post_Ex_m_diff*100)) +
        annotate("text", label = "Black vertical line  =\n mean outcome \n difference (3.2%) \nincreased with PCI", x = 5.5, y = .15, color = "black",  hjust = 0) +
        annotate("text", label = "Grey AUC = probability (94%) \n PCI > CABG outcome by > 1%", x = 4.2, y = .24, color = "black",  hjust = 0) +
         annotate("text", label = "Grey + yellow AUC =\n probability (99%) \n PCI > CABG outcome", x = -1, y = .22, color = "black", hjust = 0) +
        theme_classic() +
        stat_function(fun = dnorm, args = list(mean(post_Ex_m_diff*100), sd(post_Ex_m_diff*100)), xlim = c(1,8), geom = "area", alpha = 0.2) +
        stat_function(fun = dnorm, args = list(mean(post_Ex_m_diff*100), sd(post_Ex_m_diff*100)), xlim = c(0,1), geom = "area", alpha = 0.2, fill = "yellow") 
fig1b

pdf("output/fig1b.pdf")
fig1b 
dev.off()

ggsave("output/fig1b.png")  
```



### EXCEL MACCE outcome difference _ Figure 1c
Bayesian probability difference based on non-informative prior

```{r}
# Excel data
Ex_r_c <- temp[1,6]
Ex_r_e <- temp[1,7]
Ex_n_c <- temp[1,8]
Ex_n_e <- temp[1,9]

set.seed(1234)
#Prior is beta(1,1)
# sampling 100,000 RV from posterior
post_Ex_r_c <- rbeta(100000, Ex_r_c + 1, Ex_n_c - Ex_r_c + 1 )
post_Ex_r_e <- rbeta(100000, Ex_r_e + 1, Ex_n_e - Ex_r_e + 1 )

# calculting posterior of differences
post_Ex_r_diff <- post_Ex_r_e - post_Ex_r_c
paste("EXCEL data alone - Differences in MACCE between PCI & CABG")
quantile(post_Ex_r_diff, probs = c(0.025, .5, 0.975))

# probability > 0
sum(post_Ex_r_diff*100 >0)/100000
# probability > 3
sum(post_Ex_r_diff*100 >3)/100000

Ex_diff_df <- data.frame(post_Ex_r_diff)
Ex_diff_df$post_Ex_r_diff <- Ex_diff_df$post_Ex_r_diff*100


# graph
fig1c <- ggplot(data.frame(x = c(0, 14)), aes(x = x)) +
        stat_function(fun = dnorm, args = list(mean(post_Ex_r_diff*100), sd(post_Ex_r_diff*100)), colour = "orange", size=2) +
        scale_x_continuous(name = "Secondary composite outcome difference (PCI - CABG) -> PCI worse") +
        scale_y_continuous(name = "Density") +
        labs(title="EXCEL PDF for secondary composite outcome (risk difference)",
             subtitle = "Non-informative prior", caption="AUC = area under the curve\n     PDF = probability density function") +
        geom_vline(xintercept=mean(post_Ex_r_diff*100)) +
        annotate("text", label = "Black vertical line  = mean outcome \n difference (6.7%) increased with PCI", x = 8, y = .18, color = "black",  hjust = 0) +
        annotate("text", label = "Grey AUC = probability (97%) \n PCI > CABG outcome by > 3%", x = 9, y = .12, color = "black",  hjust = 0) +
         annotate("text", label = "Grey + yellow AUC = \nprobability (99.9%) \n PCI > CABG outcome", x = 0.5, y = .16, color = "black",  hjust = 0) +
        theme_classic() +
        stat_function(fun = dnorm, args = list(mean(post_Ex_r_diff*100), sd(post_Ex_r_diff*100)), xlim = c(3,14), geom = "area", alpha = 0.2) +
        stat_function(fun = dnorm, args = list(mean(post_Ex_r_diff*100), sd(post_Ex_r_diff*100)), xlim = c(0,3), geom = "area", alpha = 0.2, fill = "yellow") 
fig1c

pdf("output/fig1c.pdf")
fig1c 
dev.off()

ggsave("output/fig1c.png")  
```

## Combining graphs

```{r fig.width=20}
library (patchwork)

fig1 <- fig1a + fig1b + fig1c
fig1

pdf("output/fig1.pdf")
fig1
dev.off()

ggsave("output/excel_fig3.png")  

## create jpeg file
jpeg(filename = "output/fig1.jpeg", width = 7480, height = 7480,
     pointsize = 8, quality = 75, bg = "white", res = 600)
fig1
dev.off()
```

## Session info

```{r}
R.version
sessionInfo()
```

## References